---
title: Project3
author: Chris He
date: "`r Sys.Date()`"
output: html_document
Date Created: Tuesday, 2024/04/09, 16:50
Date Modified: Monday, 2024/04/15, 00:56
class: MTH362 Statistical Modeling
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE) 
```

## Introduction

The dataset comes from FiveThirtyEight, a statistics focused news website that reported on sports, politics, and science.It contains the information of 233 ads from 10 brands that aired the most ad spots in 21 Superbowls ranging from 2000 to 2021 on the social media platform known as YouTube. The research question that is being investigated is "What factors may lead to a popular commercial during the super bowl? Given that the Superbowl is one of the most watched events each year with the most expensive ad slots, it makes sense to investigate factors that impact popularity and perception of money spent on ads during the Superbowl.

```{r}
library(tidyverse)
library(tidyr)
library(leaps)
youtube <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-02/youtube.csv')
head(youtube)
youtube$popular<- with(youtube,ifelse(view_count>200000,1,0))
```

The dataset contains 247 observations of 25 variables. A derived boolean variable called `popular` was created using `view_count`, using an arbitrarily selected limit of 200,000 as a marker for a popular commercial video. As commercials ran during the Superbowl are expensive, it makes sense to set the standards of `popular` higher rather than lower. `popular` will be the response variable investigated in this project.

## EDA

```{r}
ggplot(data=youtube,aes(x=popular))+geom_bar()
```

It appears that most of the commercials shown during the Superbowl did not gain popularity. Approximately 50 videos did, while upwards of 175 did not.

```{r}
ggplot(data=youtube, aes(x=popular))+geom_bar()+facet_wrap(~show_product_quickly)
ggplot(data=youtube, aes(x=popular))+geom_bar()+facet_wrap(~use_sex)
```

Showing the product quickly seems to increase the popularity of the video more than showing it for a longer period of time. Although non popular videos double, popular videos almost triple. People don't like it when ads show the product too heavily.

Using sex seems to be relatively unimpactful to the chances of popularity.

## Model

As the chosen response variable is binary, a logistic regression was chosen to represent the odds of a video being popular. The explanatory variables selected for the model were funny, show_product_quickly, patriotic, celebrity, danger, animals, use_sex, like_count, dislike_count.

```{r}
logit.reg<- glm(popular ~ funny+ show_product_quickly+patriotic+celebrity+danger+animals+use_sex+like_count+dislike_count,data=youtube)
summary(logit.reg)
```

The variables deemed significant with a p value below 0.05 were like_count and show_product_quicklyTRUE. patrioticTRUE was close enough at 0.065 to be incorporated in the model

The resulting model was $logit(p)=0.1290\text{show_product_quicklyTRUE}+0.1487patrioticTRUE+8.211e^{-06}\text{like_count}$

If a commercial shows their product quickly, odds of popularity increase by $e^{0.1290}=1.13769012417$

If a commercial is patriotic, odds of popularity increase by $e^{0.1487}=1.16032483954$

Per a unit increase in like count, the odds of popularity increase by $e^{8.211e^{-06}}=1.00000821103$

## Analysis

```{r}
ggplot(data=youtube,aes(x=like_count,y=popular))+geom_jitter(alpha=0.5, height=0.1) + geom_smooth(method='glm', method.args=list(family='binomial'), se=TRUE) + labs(x='like_count', y='Popular?: 0 = No, 1 = Yes')
```

It appears once like counts hits a certain point, the video will be popular

```{r}
arm::binnedplot( x=logit.reg$fitted, y=logit.reg$resid, xlab="Predicted Probabilities", main="Bin Resid vs. Pred Prob", col.int = FALSE)
```

The residuals appear evenly distributed over and under the line and does not feature any patterns.

```{r}
1-pchisq(logit.reg$deviance, logit.reg$df.residual)
```

With a value of 1 from the chi squared test, it suggests that we cannot reject the null hypothesis of the logistic regression being a good fit for the data. The logistic regression seems like an appropriate model for this dataset.

## Conclusion

With a strong logistic model that follows logistic regression assumptions from the ads from 2000 to 2021, brands that are looking to make a popular commercial for Superbowls should look to show their product quickly, make it patriotic, and to have a likeable ad on Youtube.
